---
title: "R Machine Learning Pilot (Day 2)"
output: html_notebook
editor_options: 
  markdown: 
    wrap: 72
---

# Recap of Part 1

In Part 1 of this workshop, we attempted to predict the value of a
single, continuous outcome variable `hourly_wage` based on a range of
input features, including age, location, and industry. Specifically, we
covered:

-   Train-test splits: partitioning the data set into a training set
    used to train our model, and a test set used to evaluate the model's
    performance.

-   The bias-variance tradeoff: the balance between the model's ability
    to capture the true underlying patterns of the training data (bias)
    and its flexibility to learn from specific instances within that
    data (variance). This tradeoff matters because it directly affects
    the generalization ability of the model to new, unseen data.

-   Pre-processing: cleaning and transforming raw data into a suitable
    format for analysis, ensuring the data is free of errors,
    inconsistencies, and irrelevant information. It often includes
    coding categorical variables and normalizing numerical values to
    reduce bias and improve model accuracy.

Today, we will apply and expound upon these principles to develop models
to predict categorical variables. This process is called
'classification.'

# Install Packages and Load Data

Before we dive into today's material, let's load our libraries
`tidymodels` and `tidyverse` libraries.

```{r install, include=FALSE}
# install.packages(c("tidyverse","tidymodels"))
library(tidymodels)
library(tidyverse)

# Prefer tidymodels functions in any case of name conflict
tidymodels::tidymodels_prefer() 

# set seed for reproducibility 
set.seed(123)
```

# Other Tasks in Supervised Machine Learning: Classification

Thus far, we've spent most of our time working through regression
problems (i.e., predicting a continuous outcome variable). Let's switch
to a new task: classification. In classification, we aim to predict one
of a group of values. For example, predicting a qualitative response is
considered classification because we assign the observation to a
category (or class).

Like regression, classification is also a supervised learning technique
because we have a set of labeled training data that we can use to build
a classifier. Now, however, we have access to different techniques
because the structure of our outcome variable is categorical.

📝 **Poll 1**: Identify which of the following are classification
problems in machine learning.

1.  An advertiser is interested in the relationship between age and the
    number of hours of YouTube consumed.

2.  A medical testing company conducts a procedure to determine whether
    a person has a cancer diagnosis.

3.  A researcher is interested in the effect of an education
    intervention on students' test scores.

4.  A software engineer is designing an algorithm to detect whether an
    email is spam or not.

5.  A political scientist wants to classify Twitter posts as positive or
    negative.

**Solution 1: (2), (4), and (5).**

# Today's Data: Voting Patterns in the 2020 Election

Now, let's load our primary data set for today's workshop: `vote2020`.
Our goal is going to be predicting whether someone voted in the 2020
election, perhaps to tailor engagement strategies toward those least
likely to vote in an upcoming election.

```{r}
vote2020 <- read.csv("../data/vote2020.csv",row.names = NULL)

# visually inspect the data frame 
summary(vote2020)
```

## Exploratory Analysis

Recall that the first step in machine learning (and most analyses) is to
explore the data we're working with to get a sense of its shape and
anticipate problems that may arise.

🥊 **Challenge 1**: Perform exploratory analyses on the `vote2020` data
set, keeping in mind that today's goal is to predict `voted`. What do
you notice about the data?

```{r}
# Load necessary libraries
vote2020 %>%
  group_by(state) %>%
  summarise(voted_percentage = mean(voted, na.rm = TRUE) * 100) %>%
  ungroup() %>%
  ggplot(aes(x = reorder(state,-voted_percentage), y = voted_percentage)) +
    geom_bar(stat = "identity", fill = "blue") +
    theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust = 1)) +  # Adjusting x-axis text +
  coord_flip() + 
  scale_y_continuous(limits=c(0,100)) + 
    labs(title = "Percentage of People Who Voted in Each State",
         x = "State",
         y = "Percentage Voted") +
    theme_minimal()
```

## Train-Test Split

Recall from Part 1 that a critical first step in machine learning is
partitioning our data into training and test sets.

🔔 **Question 1**: Using the entire `vote2020` data set for model
training resulted in impressively high accuracy during evaluation.
However, upon deploying this model to predict voter turnout for an
upcoming election, the predictions significantly diverged from actual
turnout, with many discrepancies in who was predicted to vote versus who
actually voted. The model's near-perfect performance in development
starkly contrasted its poor real-world prediction outcomes, indicating a
potential oversight in the model training and evaluation process.

**Answer 1**: The key issue is that the model was never tested on unseen
data. By using the entire data set for training, the model essentially
"memorized" the data, including any noise or patterns specific to that
set of individuals. This led to overfitting, where the model excelled at
predicting the training data but failed to generalize to new, unseen
data. Since the model's performance was only evaluated on data it had
already seen, its apparent accuracy was misleading, giving a false sense
of confidence in its predictive capabilities.

Now that we have a better sense of the data, let's go ahead and split
`vote2020` into training and test sets:

```{r}
# Perform splits
vote2020$voted <- as.factor(vote2020$voted)
vote_split <- initial_split(vote2020, prop = 0.75)
vote_train <- training(vote_split)
vote_test <- testing(vote_split)
```

## Pre-Processing

In our example from Part 1, we prepared our data for analysis by
recoding categorical variables and normalizing numeric ones using
`step_dummy(all_nominal_predictors())` and `step_normalize`,
respectively.

In that example, we also dropped rows that had any missing values across
variables. Let's try another example, in which we don't *omit* samples
that have missing values, but instead perform *imputation*, in which we
replace those missing values according to certain criteria. There are
various kinds of imputation, including:

-   For example, whenever we have a missing value for the `occupation`,
    we can replace that missing value with the most common occupation.
    This is called *mode imputation*.

-   Or, we could replace a missing numerical predictor (e.g., age) using
    the median across all the samples. This is called *median
    imputation*.

There are other ways to impute, but these are good starting points. The
way to perform imputation in a `recipe` is via the `step_impute_*`
functions.

🥊 **Challenge 2**: Using the same logic from Part 1, create your own
recipe called `voterecipe` that lays out the steps for pre-processing
the data. Instead of dropping rows with missing values, use
`step_impute_median` to impute numeric variables and `step_impute_mode`
to impute categorical variables.

```{r}
voterecipe <- 
  recipe(voted ~ ., data = vote_train) %>%
  step_impute_median(all_numeric_predictors()) %>%
  step_impute_mode(all_nominal_predictors()) %>%
  step_dummy(all_nominal_predictors()) %>% 
  step_normalize(all_of(c("age")))
```

This pre-processing will allow us to take advantage of samples with
missing data, even if it comes at a little cost to accuracy. Imputation
is often a necessary step, since it's common to have missing data.

## Developing and Evaluating Models

Now that we've performed exploratory analyses, pre-processing, and a
train-test split, let's go ahead and train our model. We will create a
classifier (i.e., a model that predicts membership in a group) with two
methods: logistic regression and decision trees.

#### Algorithm 1: Logistic Regression

Machine learning practitioners often recommend logistic regression as a
starting model when predicting a binary outcome or probability. For
example, if we are estimating the relationship between mortality and
income we can estimate the probability of mortality given a change in
income.

We write this as $P[M|I]$ and the values will range between 0 and 1. We
can make a prediction for any given value of income on the mortality
outcome. Normally, we establish a threshold for prediction. For example,
we might predict death where $P[M|i] > 0.5$.

Logistic regression is a generalized linear model where we model the
probability function using the logistic function. Define the general
function as $p(Y= 1|X)$. Then the model of interest is:

$$p(Y = 1|\textbf{X}) = \frac{e^{\textbf{X}\beta}}{1 + e^{\textbf{X}\beta}}$$
Here the bold X indicates a vector of features and the $\beta$
coefficient represents a vector of coefficients.

Here's how to fit classification problems in `tidymodels` using logistic
regression.

Let's go ahead and perform the training and test splits for the `vote`
data set:

```{r}
# Perform splits
vote2020$voted <- as.factor(vote2020$voted)
class_split <- initial_split(vote2020, prop = 0.75)
class_train <- training(class_split)
class_test <- testing(class_split)
```

In `tidymodels`, creating a logistic regression follows the exact same
procedure as a linear regression. This time, however, we will use
`logistic_reg()` to initiate the function. Let's create the model:

```{r}
# Create model
logistic_model <- logistic_reg(mode = "classification")
```

Next, let's create a `recipe` for pre-processing:

```{r}
# Recipe for classifying defaults
class_recipe <- recipe(voted ~ ., data = class_train) %>%
  step_impute_median(all_numeric_predictors()) %>%
  step_impute_mode(all_nominal_predictors()) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_normalize(all_of(c("age")))
  
```

Now, we create our workflow, and fit the model:

```{r}
# Create workflow 
class_wflow <- workflow() %>%
  add_recipe(class_recipe) %>%
  add_model(logistic_model)

# Fit the workflow
class_fit <- class_wflow %>% 
  fit(data = class_train)
```

Finally, let's use augment from the `yardstick` package to obtain the
predictions, and take a look at them:

```{r}
predictions <- augment(class_fit, new_data = class_test)
predictions
```

🔔 **Question 2**: Notice that new three columns have appeared in our
data set. What do these values mean? What are they telling us?

**Answer 2**: `.pred_class` corresponds to the value our algorithm is
predicting, i.e., whether the person voted in the election or not. This
was determined using the two variables `.pred_0` and `.pred_1`, which
represent the predictions from the logistic regression. If membership in
group 1 (i.e., voted = 1) is greater than .5, then `.pred_class` will
take a value of 1.

To evaluate the model, we'll use the `accuracy` function again:

```{r}
accuracy(predictions, truth = voted, estimate = .pred_class, type="class")
```

We predicted the likelihood that someone voted in the last election with
an accuracy of 81% - not bad! We can also create what's called a
'confusion matrix' to see how well our model did at predicting each
class.

```{r}
predictions %>%
  conf_mat(truth = voted, estimate = .pred_class)
```

💡 **Tip**: While we used logistic regression to predict a binary
outcome `voted`, recall that this prediction is based on a range of
likelihoods from 0 to 1. These predictions can be useful to supplement
causal inferences methods like propensity score matching that require
precise likelihood estimates without having to explain why.

But, there are many other options in our machine learning tool kit to get a better prediction. 

#### Hyperparameter Tuning and Regularization 

Notice that until now, we have used the default settings for the linear
and logistic regressions we have run. Looking at the documentation for logistic regression, `?logistic_reg()`, we see that there are many arguments that we left blank when we initialized our model above. These arguments, also known
as 'parameters', correspond to statistical choices about how the model
should operate or be structured. 

Some of these hyperparameters include 'engine', 'penalty', and 'mixture': 

* Engine: Different engines can implement regression through various algorithms or computational approaches. When you specify an engine for logistic regression in R, you're choosing the particular set of algorithms and optimizations that will be used to train your model.

- Penalty: "penalty" refers to a regularization technique used to prevent overfitting by discouraging overly complex models. It does this by adding a penalty to the loss function for large coefficients. Common penalties are L1 (Lasso), which can shrink some coefficients to zero (thus performing feature selection), and L2 (Ridge), which shrinks all coefficients toward zero but typically doesn't set any to exactly zero. The penalty helps in creating simpler, more generalizable models that perform better on unseen data by prioritizing the most influential features and reducing the model's sensitivity to the training data's noise.

For example, we can add a 'penalty' on the size of the coefficients of a
model and reduce the likelihood of overfitting. Lasso, ridge, and
elastic net are different types of penalties that greatly reduce or
shrink to zero coefficients on variables that are picking up on a lot of
noise. The broad technique of reducing overfitting is called
'regularization.'

🥊 **Challenge 3**: We have reproduced the original code from above that trained a basic classifier on our voting data set without changing any of the hyperparameters, as well as the code that obtains predictions. Re-run this code several times, but each time change the hyperparameters in the model specification. For mixture, select values between 0 and 1 inclusive; for penalty, include a non-negative number; and for engine, select either "glmnet" or "glm". How does the accuracy change? 

```{r}
chal3_logistic_model <- logistic_reg(mode = "classification", 
                                     mixture=0, # ranges between 0 and 1
                                     penalty = 0, # greater >= 0 
                                     engine="glmnet") 

# Create workflow 
chal3_wflow <- workflow() %>%
  add_recipe(class_recipe) %>%
  add_model(chal3_logistic_model)

# Fit the workflow
chal3_fit <- chal3_wflow %>% 
  fit(data = class_train)

# obtain predictions 
chal3_predictions <- augment(chal3_fit, new_data = class_test)
accuracy(chal3_predictions, truth = voted, estimate = .pred_class, type="class")
```

Cue hyperparameter tuning! Hyperparameter tuning is crucial in machine
learning as it directly impacts the performance and effectiveness of
models. By fine-tuning hyperparameters, practitioners can optimize
models to achieve higher accuracy, better generalize to unseen data, and
prevent issues like overfitting or underfitting. This process allows for
the customization of models to specific datasets and objectives,
enabling the discovery of the best configuration for a given problem.






APPENDIX: ADVANCED TOPICS

# Random Forests

The last algorithm we'll look at is the random forest, which is an
**ensemble** of decision trees. Ensembling is a common machine learning
approach in which we combine many smaller models into one final
predictions.

A decision tree is a method of stratifying the feature space into a
number of simple regions. Usually, trees use the mean or mode dependent
variable value for the training data in the region to which it belongs
(James *et al.* 2021). Trees can be applied to both regression or
classification problems. To build a single tree, we:

1.  Divide the feature space into distinct regions.
2.  For every observation that falls into a distinct region ($R_j$) we
    make the same prediction.

Trees by themselves turn out to not be very good at prediction compared
to other supervised learning model. However, combine them into a *random
forest* and they become very useful, especially if the relationship is
nonlinear and complex. A random forest (Breiman 2001) is an example of
an ensemble method, which is an approach that combines many simple
models in order to get a single powerful prediction model.

In a random forest, we build a number of decision trees on bootstrapped
training samples. Each tree is grown independently on random samples of
the observations, a process known as *bagging.*

When building the trees, each time a split occurs, we take a random
sample of features to be split candidates instead of the full set of
features. By choosing a random subset each time, a random forest
decorrelates the individual decision trees. This makes the resulting
average less variable.

Let's try fitting a random forest using the `rand_forest` model:

```{r}
# Create model specification 
rf_model <- rand_forest(
  mtry = tune(),
  min_n = tune(),
  trees = 50) %>%
  set_mode("classification") %>%
  set_engine("ranger") 
# Create recipe
rf_recipe <- recipe(union ~ ., data = class_train) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_normalize(all_predictors())
# Create workflow
rf_wflow <- workflow() %>%
  add_recipe(rf_recipe) %>%
  add_model(rf_model)
```

Random forests are known for having many hyperparameters than need
tuning. In this case, we need to tune two: `mtry` and `min_n`. Let's
create the `rsample` object for using 3 folds. Then, we'll tune the
grid. This time, we'll just specify `grid = 10` to indicate we want 10
hyperparameter configurations. The `dials` package, in this case, will
automatically choose those values for us.

```{r}
# Use cross-validation to tune our model appropriately 
trees_folds <- vfold_cv(class_train, v = 3)

# Find appropriate tuning parameters
set.seed(345)
rf_fit <- tune_grid(
  object = rf_wflow, 
  resamples = trees_folds, 
  grid = 10)
```

We're going to use a new metric this time called the ROC-AUC. This is
the Area Under the Curve for the Receiver-Operator Characteristic. The
main takeaway for this metric is that it tells us how good our
probability predictions are, rather than just the predictions. In
classification settings, it's often preferable to the accuracy. It
varies from 0 to 1, where larger is better:

```{r}
# View the ROC metric across models 
rf_fit %>% 
  collect_metrics() %>%
  filter(.metric == "roc_auc") %>%
  select(mean, min_n, mtry)
```

Now, let's finalize the workflow:

```{r}
# Select the best RF model 
best_rf_auc <- select_best(rf_fit, "roc_auc")

final_rf_wflow <- rf_wflow %>%
  finalize_workflow(parameters = best_rf_auc) %>%
  fit(data = class_train)
```

And evaluate the ROC AUC:

```{r}
final_rf_wflow %>%
  augment(new_data = class_test) %>%
  roc_auc(truth = default, estimate = .pred_No)
```

That's rather good ROC performance!

# Overview

Congratulations, you've made it! We covered the basics of supervised
machine learning in `tidymodels` in this workshop. However, there's much
more to explore. The best way to keep pushing forward is to choose a
problem to study, and refer to the documentation when you need help. The
website Kaggle has an abundance of good data science problems to work on
if you need help choosing a task!

# Other classification problems

-   sentiment analysis with natural language processing--positive,
    negative, or neutral problems.
-   
